### 2015-08-04 LensingPaper_draft_v1.01.tex
Restructured paper and wrote a first (too long) version of abstract.

\documentclass[useAMS,usenatbib]{mn2e}

%_____GENERAL PACKAGES_____
\usepackage{amsmath}	%for \text{} in math mode
\usepackage{amssymb}	%for symbols

%_____FIGURES_____
\usepackage{graphicx}	%for including images
\usepackage{float}	%for [Hhtb] etc of figures
\usepackage{grffile}    %for using underscores in figure filenames
\usepackage{subcaption} %for having sub-captions in figures with multiple panels
\usepackage{placeins} %for \FloatBarrier, which forces all figures to be printed before continuing

%_____SYMBOLS_____
\usepackage{ mathrsfs } %for slanted letters in math mode

%_____FORMAT_____
\setcounter{secnumdepth}{2}	%depth of section numbering

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title[A spiral galaxy's mass distribution uncovered]{A spiral galaxy's mass distribution uncovered through lensing and dynamics}
\author[W. Trick, G. van de Ven and A. Dutton]{W. Trick$^{1}$\thanks{E-mail:
trick@mpia.de}, G. van de Ven$^{1}$ and A. Dutton$^{1}$\\
$^{1}$Max-Planck-Institute for Astronomy, K\"{o}nigstuhl 17, 69117 Heidelberg, Germany}
\begin{document}

\date{Accepted ????. Received ????; in original form ????}

\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2014}

\maketitle

\label{firstpage}

%--------------------------------------------
%Abstract
\begin{abstract}
\include{docs/abstract_v1.01}
\end{abstract}
%--------------------------------------------

\begin{keywords}
blabla -- blabla: bla.
\end{keywords}

%--------------------------------------------
%Introduction
\include{docs/introduction_v1.01}
%--------------------------------------------

%------------------------
\include{docs/data_v1.01}
%------------------------

%------------------------
\section{Modelling}
\include{docs/model_MGE_v1.01}
\include{docs/model_lensing_v1.01}
\include{docs/model_fitting_v1.01}
\include{docs/model_JAM_v1.01}
%------------------------

%------------------------
\section{Results}
\include{docs/results_photometry_v1.01}
\include{docs/results_lensing_v1.01}
\include{docs/results_JAMwithSB_v1.01}
\include{docs/results_JAMwithNFW_v1.01}
%------------------------

%------------------------
\section{Discussion and Conclusion}
\include{docs/discussion_merger_v1.01}
\include{docs/summary_v1.01}
%------------------------

%---------------------------------------------------------------------------
\bibliographystyle{mn2e}
\bibliography{literaturelist}
%---------------------------------------------------------------------------

\label{lastpage}

\end{document}