### 2015-08-04 docs/data_v1.01.tex

\section{Data}

[TO DO]

\begin{itemize}
\item Hubble Space teleskope (HST) imaging by \cite{SWELLSI} in two filters (F450W amd F814W)
\item I filter: for surface brightness distribution of J1331's bulge for Jeans modelling
\item ??? filter: to identify bluish lensing images
\end{itemize}