### 2015-08-04 docs/discussion_merger_v1.01.tex

\subsection{Does J1331 have a Merger History?}

[TO DO]