### 2015-08-04 docs/model_MGE_v1.01.tex

\subsection{Multi-Gaussian Expansion Formalism}

[TO DO]