### 2015-11-13 LensingPaper_draft_v3.01.tex
Included corrections in abstract and introduction.

\documentclass[useAMS,usenatbib]{mnras}

%_____GENERAL PACKAGES_____
\usepackage{amsmath}	%for \text{} in math mode
\usepackage{amssymb}	%for symbols

%_____FIGURES_____
\usepackage{graphicx}	%for including images
\usepackage{float}	%for [Hhtb] etc of figures
\usepackage{grffile}    %for using underscores in figure filenames
\usepackage{subcaption} %for having sub-captions in figures with multiple panels
\usepackage{placeins} %for \FloatBarrier, which forces all figures to be printed before continuing

%_____SYMBOLS_____
\usepackage{ mathrsfs } %for slanted letters in math mode

%_____FORMAT_____
\setcounter{secnumdepth}{2}	%depth of section numbering
\usepackage{hyperref} %ready-to-lick referencing

%_____MACROS______
\newcommand*\diff{\mathop{}\!\mathrm{d}}
\newcommand*\Diff[1]{\mathop{}\!\mathrm{d^#1}}
%\newcommand{\vect}[1]{\boldsymbol{#1}} % Uncomment for BOLD vectors.
\newcommand{\vect}[1]{\vec{#1}} % Uncomment for ARROW vectors.
\usepackage[usenames,dvipsnames]{xcolor}
\newcommand{\Wilma}[1]{\textcolor{Magenta}{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title[A spiral galaxy's mass distribution uncovered]{A spiral galaxy's mass distribution uncovered through lensing and dynamics}
\author[W. H. Trick, G. van de Ven and A. A. Dutton]{Wilma H. Trick$^{1}$\thanks{E-mail:
trick@mpia.de}, Glenn van de Ven$^{1}$ and Aaron A. Dutton$^{1}$\\
$^{1}$Max-Planck-Institute for Astronomy, K\"{o}nigstuhl 17, 69117 Heidelberg, Germany}
\begin{document}

\date{Accepted ???. Received ???; in original form ???}

\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2015}

\maketitle

\label{firstpage}

%--------------------------------------------
%Abstract
\begin{abstract}
\input{docs/abstract_v3.01}
\end{abstract}
%--------------------------------------------

\begin{keywords}
blabla -- blabla: bla. \Wilma{[TO DO: Figure out proper MNRAS keywords.]}
\end{keywords}

%--------------------------------------------
%Introduction
\input{docs/introduction_v3.01}
%--------------------------------------------

%------------------------
\section{Modelling} \label{sec:Modelling}
\input{docs/model_MGE_v3.01}
\input{docs/model_lensing_v3.01}
\input{docs/model_JAM_v3.01}
%------------------------

%------------------------
\input{docs/data_v3.01}
%------------------------

%------------------------
\section{Results} \label{sec:Results}
\input{docs/results_photometry_v3.01}
\input{docs/results_lensing_v3.01}
\input{docs/results_JAMwithSB_v3.01}
\input{docs/results_JAMwithNFW_v3.01}
%------------------------

%------------------------
\section{Discussion and Conclusion} \label{sec:Discussion}
\input{docs/discussion_ML_v3.01}
\input{docs/discussion_merger_v3.01}
\input{docs/summary_v3.01}
%------------------------

%---------------------------------------------------------------------------
\bibliographystyle{mnras}
\bibliography{literaturelist}
%---------------------------------------------------------------------------

\label{lastpage}

\end{document}