### 2015-11-13 docs/summary_v3.01.tex

\subsection{Summary}

\Wilma{[TO DO: Ask Glenn and Aaron how much of the above should find its way into the summary.]}